Evaluating Cultural Transmission in Bronze Age burial rites of Central, Northern and North-western Europe using radiocarbon data

### 2019-03-17 article/03_Results.Rmd
detached boring parts of the discussion and moved them to a new main chapter "supplementary remarks"
removed simulation and everything related to it from the text: careful rereading will be necessary

## Timeseries {#time-series-results}

@@

The simulation as presented in Figure \ref{simulation} was implemented to contextualize this observation of low correlation between spatial and cultural distance. Some general remarks: In case of equal interaction intensity $I$ of separate populations, the correlation of cultural and spatial distance is weak (Figure \ref{simulation} A vs. C vs. E). It appears to be normally distributed around zero for many simulation runs, although individual runs can still yield high positive and negative correlation for certain time slots. The fluctuation of the correlation within one run can be strong and seems to increase for greater values of $m_i$. If the interaction matrix $I$ is predefined to represent the spatial distance matrix (Figure \ref{simulation} BDF), then the mean correlation per time slot increases. The correlation increases as well with the degree of interaction $m_i$ (Figure \ref{simulation} D vs. F). Counter-intuitively the fluctuation of the correlation is also very high here (Figure \ref{simulation} E vs. F). The correlation development observed for the variables *burial type* and *burial construction* fits to simulation runs in all presets, but becomes less likely if we assume $I$ to represent spatial distance and a high degree of interaction $m_i$ (Figure \ref{simulation} A vs. F). That's mostly because of the relatively strong negative correlation of cultural and spatial distance in the Middle and Late Bronze Age (1400 to 1000 calBC). Despite this tendency, the real-world correlation time series are still within the margins of potential toy model events. Therefore it is not possible to rule out the influence of spatial proximity as an important factor for the Urnfield burial rite diffusion. On the contrary, the simulation results challenge an overly simplistic interpretation of Table \ref{tab:distanceandcorrelationtable*}.

>>>>

## Time series {#time-series-results}

@@

### 2019-03-17 article/04_Discussion.Rmd
removed simulation and everything related to it from the text: careful rereading will be necessary

Concerning the question of cultural distance, the expectation based on Cultural transmission theory was a clear correlation of the "cultural" SED distances for both variables *burial type* and *burial construction* with spatial distance. The simulation, though with this implementation and parametrisation for sure rather qualitatively than quantitatively, shows how an increase in the degree of interaction should translate to a higher degree of cultural similarity across human populations (Figure \ref{simulation}). And as the European Bronze Age is usually depicted as a period of trade, travel and exchange [e.g. @kristiansenRiseBronzeAge2005; @sherrattWhatWouldBronzeAge1993] there was no reason to expect something else than Tobler's famous First Law of Geography: "everything is related to everything else, but near things are more related than distant things." [@toblerComputerMovieSimulating1970].

@@

3. The toy simulation based on @neimanStylisticVariationEvolutionary1995 suggests that even under the assumption of strong transmission between directly neighbouring groups, random and emergent effects can still cause 	fluctuation and temporary lack of correlation in human populations.

4. All observations of cultural distance in this paper are based on only the two variables *burial type* and *burial construction*. This is by far not enough for any overarching conclusions about the general state of society in the Bronze Age.

>>>>

Concerning the question of cultural distance, the expectation based on Cultural transmission theory was a clear correlation of the "cultural" SED distances for both variables *burial type* and *burial construction* with spatial distance. As the European Bronze Age is usually depicted as a period of trade, travel and exchange [e.g. @kristiansenRiseBronzeAge2005; @sherrattWhatWouldBronzeAge1993] there was no reason to expect something else than Tobler's famous First Law of Geography: "everything is related to everything else, but near things are more related than distant things." [@toblerComputerMovieSimulating1970].

@@

3. All observations of cultural distance in this paper are based on only the two variables *burial type* and *burial construction*. This is by far not enough for any overarching conclusions about the general state of society in the Bronze Age.
